---
title: "Choropleth map"
description: "An sf object drawn directly with mark_sf() and a proper map projection via coord_sf()."
categories: [Spatial]
---

```{r}
#| include: false
source("_helper.R")
```

vellumplot reads simple-features geometry directly. `mark_sf()` draws the
polygons and `coord_sf()` handles the projection and aspect ratio.

For a version you can hover, select, and zoom, see the
[interactive choropleth](choropleth-interactive.qmd).

```{r}
nc <- sf::st_read(system.file("shape/nc.shp", package = "sf"), quiet = TRUE)

p <- vplot(nc, width = 8, height = 3.6) |>
  mark_sf(fill = BIR74) |>
  coord_sf() |>
  labs(title = "Births by county, North Carolina (1974)", fill = "births")
```

```{r}
#| echo: false
vg(p, "choropleth")
```

## Dissolving the borders between equal values

Bin the values and a per-feature choropleth starts to lie about its own geometry:
neighbouring counties in the same class are drawn as separate polygons, so a
border is stroked twice down the middle of what should read as one region — a
visible seam, and in PDF an exact one.

`mark_sf(merge = TRUE)` dissolves adjacent features that share a fill into a
single region. The union is real boolean geometry (`vellum::vl_path_op()`), holes
and multipart features preserved, so each region comes out as one crisp path
rather than a stack of alpha-composited neighbours.

```{r}
nc$class <- cut(nc$BIR74, breaks = c(0, 1000, 2000, 4000, Inf),
                labels = c("under 1k", "1k–2k", "2k–4k", "over 4k"))

p_merged <- vplot(nc, width = 8, height = 3.6) |>
  mark_sf(fill = class, merge = TRUE) |>
  coord_sf() |>
  labs(title = "Birth classes, merged into regions", fill = "births")
```

```{r}
#| echo: false
vg(p_merged, "choropleth-merged")
```

Compare it with the same classes drawn feature by feature — same colours, but every
county boundary still there, including the ones between counties of the same class:

```{r}
p_unmerged <- vplot(nc, width = 8, height = 3.6) |>
  mark_sf(fill = class, color = "white", linewidth = 0.2) |>
  coord_sf() |>
  labs(title = "The same classes, one path per county", fill = "births")
```

```{r}
#| echo: false
vg(p_unmerged, "choropleth-unmerged")
```

`merge` is ignored on an interactive layer, and necessarily so: merging throws away
the per-feature keys a widget needs to know which county you hovered.

For the cartographic furniture — a scale bar, a north arrow, a graticule — see
[Scale bar, north arrow, graticule](map-decorations.qmd).
