---
title: "Chord diagram"
description: "Directed flows between a handful of nodes, wrapped onto a circle with an out/in split and self-loops."
categories: [Flows]
---

```{r}
#| include: false
source("_helper.R")
```

`vchord()` wraps a weighted flow onto a circle: one arc per node, sized by the
node's total incident weight, and one ribbon per flow joining the source arc to
the target arc through the centre. It takes a `from`/`to`/`value` flow list or a
square flow matrix (row = from, column = to).

Here each cell is the number of customers who moved from one carrier to another
over a year; the diagonal is those who stayed.

```{r}
switch <- matrix(
  c(
    40, 12, 8, 5,
    10, 55, 6, 9,
    7, 8, 38, 11,
    6, 10, 9, 44
  ),
  nrow = 4, byrow = TRUE,
  dimnames = list(
    c("Aircel", "Boost", "Cellcom", "Ditto"),
    c("Aircel", "Boost", "Cellcom", "Ditto")
  )
)
p <- vchord(switch)
```

```{r}
#| echo: false
vg(p, "chord-switch", dpi = 150)
```

Flows are directed, so each node's arc is split into an outgoing block and an
incoming block. `Aircel -> Boost` and `Boost -> Aircel` become separate ribbons,
and each carrier's retained customers (`Aircel -> Aircel`) loop as a self-ribbon
within its own sector. Ribbons take their source node's colour; pass
`link_color = "target"` to colour by destination instead, and `sort = "value"`
to order the sectors by weight.

```{r}
p <- vchord(switch, sort = "value", link_color = "target")
```

```{r}
#| echo: false
vg(p, "chord-target", dpi = 150)
```
