---
title: "Neon glow and gradient fill"
description: "A high-contrast cyberpunk theme with a glowing line and a gradient area fill."
categories: [Effects]
---

```{r}
#| include: false
source("_helper.R")
```

Layer effects and gradient fills are first-class. A `linear_gradient()` fades
the area to transparent, `glow()` haloes the line, and `theme_cyberpunk()` sets
the near-black canvas and neon palette.

The halo is a real Gaussian blur in the engine (`vellum::vl_viewport(blur = )`):
one softened layer composited under the line, rather than a stack of widened
low-opacity copies faking a soft edge. It is cheaper and smoother, and it lifts the
restriction that came with the old approach — glowing text.

```{r}
set.seed(1)
area <- data.frame(x = 1:80, y = 20 + cumsum(rnorm(80, 0.05, 1)))

p <- vplot(area) |>
  mark_area(
    x = x, y = y,
    fill = linear_gradient(c("#08F7FE", "#08F7FE00"),
                           x1 = 0, y1 = 1, x2 = 0, y2 = 0)
  ) |>
  mark_line(x = x, y = y, color = "#08F7FE", effects = list(glow(size = 5))) |>
  labs(title = "Gradient fill under a neon line") |>
  theme_cyberpunk()
```

```{r}
#| echo: false
vg(p, "cyberpunk")
```

## Text glows too

`mark_text()` and `mark_label()` take an `effects` argument, and `glow()` and
`shadow()` work on them — an annotation can sit in the same neon register as the
line it is annotating instead of being the one flat element on a glowing canvas.

```{r}
peak <- area[which.max(area$y), ]
peak$label <- "peak"
peak$x <- peak$x - 9
peak$y <- peak$y - 2

p2 <- vplot(area, width = 7, height = 4) |>
  mark_area(
    x = x, y = y,
    fill = linear_gradient(c("#FE53BB", "#FE53BB00"),
                           x1 = 0, y1 = 1, x2 = 0, y2 = 0)
  ) |>
  mark_line(x = x, y = y, color = "#FE53BB", effects = list(glow(size = 5))) |>
  mark_text(x = x, y = y, label = label, size = 22, color = "#F5D300",
            effects = list(glow(size = 8)), data = peak) |>
  labs(title = "A glowing annotation") |>
  theme_cyberpunk()
```

```{r}
#| echo: false
vg(p2, "cyberpunk-text")
```

A *sharp* `outline()` still needs a glyph outline the text primitive cannot stroke,
so that one is unavailable on text; `glow()` and `shadow()`, being blurs, are not.
