---
title: "Dendrogram"
description: "A hierarchical clustering drawn straight from hclust(), cut into coloured groups."
categories: [Hierarchies]
---

```{r}
#| include: false
source("_helper.R")
```

`vdendrogram()` draws a base `hclust` (or `dendrogram`) as a node-link tree:
leaves along one edge, each merge at its height, joined by right-angle brackets.
It coerces the clustering to `vgraph()`'s node and edge tables, so the result is
an ordinary plot spec you can theme and render like any other.

Pass `k` to cut the tree with `cutree()` and colour the branches and leaf labels
by cluster. Branches above the cut, where the clusters join, stay grey.

```{r}
hc <- hclust(dist(scale(mtcars)), method = "ward.D2")
p <- vdendrogram(hc, k = 4)
```

```{r}
#| echo: false
vg(p, "dendrogram-cut", dpi = 150)
```

With 32 leaves the labels are tight along the bottom. `direction = "right"`
grows the tree sideways and lays the labels out horizontally, which reads better
for a long list of leaves.

```{r}
p <- vdendrogram(hc, k = 4, direction = "right")
```

```{r}
#| echo: false
vg(p, "dendrogram-right", dpi = 150)
```

For finer control, skip the preset. `vgraph()` takes the clustering directly, and
`layout = "dendrogram"` reads the merge heights; you then add the marks yourself.
`layout = "unrooted"` (from `graphlayouts`) instead spreads the tree across the
plane the way phylogenies are drawn.
