---
title: "Dendrogram"
description: "A hierarchical clustering drawn straight from hclust(), with the tree cut into coloured groups."
categories: [Hierarchies]
---

```{r}
#| include: false
source("_helper.R")
```

`vdendrogram()` takes a base `hclust` (or `dendrogram`) and draws it as a
node-link tree: leaves along the bottom, each merge at its height, joined by
right-angle brackets. It coerces the clustering to `vgraph()`'s node/edge tables
under the hood, so the result is an ordinary plot spec.

```{r}
hc <- hclust(dist(scale(mtcars)), method = "ward.D2")
```

`k` cuts the tree with `cutree()` and colours the branches and leaf labels by
cluster. Branches above the cut, where separate clusters join, stay grey.

```{r}
#| echo: false
vg(vdendrogram(hc, k = 4), "dendrogram-cut", dpi = 150)
```

Pass `direction = "right"` to grow the tree sideways, which gives the labels more
room when there are many leaves.

```{r}
#| echo: false
vg(vdendrogram(hc, k = 4, direction = "right"), "dendrogram-right", dpi = 150)
```

For the general case, skip the preset: `vgraph()` accepts the clustering directly
and `layout = "dendrogram"` reads the merge heights, leaving the marks to you.
`layout = "unrooted"` (from `graphlayouts`) instead spreads an unrooted tree
across the plane, the way phylogenies are usually shown.
