---
title: "Flow map"
description: "A one-to-many flow map: migration into California fans out along merging branches sized by volume."
categories: [Networks]
---

```{r}
#| include: false
source("_helper.R")
```

A flow map shows one source reaching many destinations, the branches merging
into trunks whose width grows with the combined flow -- the idiom Minard used
for Napoleon's march and cartographers use for migration. `mark_flow_map()`
takes a one-to-many (star) graph laid out at fixed coordinates, the name of the
`root`, and a per-edge `weight`; the geometry comes from the
[edgebundle](https://github.com/schochastics/edgebundle) package.

`edgebundle::cali2010` records migration into California from every other state
in 2010. Placing each state at its geographic centre and rooting the flow at
California draws the streams converging on the west coast, each trunk sized by
how many people moved along it.

```{r}
g <- edgebundle::cali2010
xy <- cbind(state.center$x, state.center$y)[
  !state.name %in% c("Alaska", "Hawaii"),
]

p <- vgraph(g, layout = xy) |>
  mark_flow_map(root = "California", weight = weight, width_range = c(0.3, 4)) |>
  mark_nodes(size = 1, fill = "grey30") |>
  labs(title = "Migration into California, 2010")
```

```{r}
#| echo: false
vg(p, "flow-map")
```

`type = "spiral"` (the default) is an angle-restricted spiral tree: planar,
crossing-free, and needing only the edgebundle package. `type = "steiner"`
builds an approximate Steiner tree instead (it additionally needs the `interp`
package). The computed flow is mapped onto `width_range`, so the busiest trunk
stays legible however lopsided the volumes are.
