---
title: "Hierarchy diagrams"
description: "One tree drawn four ways from the same parent list: sunburst, treemap, circle-pack, and icicle, switched with a single argument."
categories: [Hierarchies]
---

```{r}
#| include: false
source("_helper.R")
```

`vhierarchy()` reads a *parent list*: one row per node carrying an `id`, its
`parent` (`NA` at the root), and a `value` for the leaves. An internal node's
value is the sum of its children, so you only supply the leaves. The `type`
argument picks the geometry, and nothing else changes, so the same tree can be
redrawn a different way by editing one word.

```{r}
budget <- data.frame(
  id = c(
    "Budget",
    "Engineering", "Sales", "Marketing",
    "Frontend", "Backend", "Infra",
    "Field", "Inside",
    "Digital", "Events"
  ),
  parent = c(
    NA,
    "Budget", "Budget", "Budget",
    "Engineering", "Engineering", "Engineering",
    "Sales", "Sales",
    "Marketing", "Marketing"
  ),
  value = c(NA, NA, NA, NA, 20, 28, 12, 18, 10, 14, 6)
)
```

The default `type = "sunburst"` maps depth to concentric rings and each node's
angular span to its share of its parent. The root is the centre; `root_label`
prints its total there.

```{r}
#| echo: false
vg(
  vhierarchy(budget, id, parent, value, show_values = TRUE, root_label = TRUE),
  "hierarchy-sunburst"
)
```

`type = "treemap"` packs the tree into nested rectangles sized by value, using a
squarified layout that keeps each rectangle close to square. Every branch keeps
its own hue, lightened one step per level, and shows as a thin frame around its
children.

```{r}
#| echo: false
vg(
  vhierarchy(budget, id, parent, value, type = "treemap", show_values = TRUE),
  "hierarchy-treemap"
)
```

`type = "circlepack"` nests circles instead of rectangles. A leaf's radius grows
with the square root of its value, and each branch is the smallest circle that
encloses its children.

```{r}
#| echo: false
vg(
  vhierarchy(budget, id, parent, value, type = "circlepack", show_values = TRUE),
  "hierarchy-circlepack"
)
```

`type = "icicle"` is the rectangular twin of the sunburst: depth becomes a stack
of adjacent bands rather than rings. Set `flow` to run the levels down, up, left,
or right.

```{r}
#| echo: false
vg(
  vhierarchy(budget, id, parent, value, type = "icicle", flow = "right"),
  "hierarchy-icicle"
)
```
