---
title: "Edge bundling"
description: "Bundled edge routing collapses a dense node-link diagram into a few legible trunks."
categories: [Networks]
---

```{r}
#| include: false
source("_helper.R")
```

Straight edges on a dense graph overlap into a hairball. `mark_edge_bundle()`
routes them as curves that merge where they run together, so the graph's
backbone shows through. It swaps in for `mark_edges()` and takes the same edge
aesthetics; the geometry comes from the
[edgebundle](https://github.com/schochastics/edgebundle) package.

Here five clusters sit apart under a stress layout, and the edges running
between them bundle into thick channels.

```{r}
set.seed(1)
g <- igraph::sample_islands(5, 18, 0.4, 6)
g <- igraph::set_vertex_attr(
  g, "grp",
  value = as.factor(igraph::cluster_louvain(g)$membership)
)

p <- vgraph(g, layout = "stress") |>
  mark_edge_bundle(type = "hammer", color = "grey30") |>
  mark_nodes(size = 2, fill = grp) |>
  labs(title = "Bundled edges between clusters", fill = "cluster")
```

```{r}
#| echo: false
vg(p, "network-bundle")
```

`type` picks the algorithm. `"force"` (the default) and `"path"` bend edges
gently, `"stub"` only tufts each endpoint, and `"mingle"` merges hierarchically;
`"divided"` splits a directed graph's trunks by direction. Bundled edges are
faint by default so overlapping trunks read as density.
