---
title: "Edge bundling"
description: "Bundled edge routing collapses 2,682 US flight routes into a few legible trunks."
categories: [Networks]
---

```{r}
#| include: false
source("_helper.R")
```

Straight edges on a dense graph overlap into a hairball. `mark_edge_bundle()`
routes them as curves that merge where they run together, so the graph's
backbone shows through. It swaps in for `mark_edges()` and takes the same edge
aesthetics; the geometry comes from the
[edgebundle](https://github.com/schochastics/edgebundle) package.

The classic demonstration is `edgebundle::us_flights`: 276 US airports and 2,682
routes between them. Laying the airports out at their real longitude/latitude
draws the outline of the country; force bundling then pulls the routes into the
transcontinental corridors.

```{r}
g <- edgebundle::us_flights
xy <- cbind(
  igraph::V(g)$longitude,
  igraph::V(g)$latitude
)

p <- vgraph(g, layout = xy) |>
  mark_edge_bundle(type = "force", color = "steelblue") |>
  mark_nodes(size = 0.2, fill = "orange") |>
  labs(title = "US domestic flight routes")
```

```{r}
#| echo: false
vg(p, "network-bundle")
```

`type` picks the algorithm. `"force"` (used here) and `"path"` bend edges toward
their neighbours, `"stub"` only tufts each endpoint, and `"mingle"` merges
hierarchically; `"divided"` splits a directed graph's trunks by direction. Pass
algorithm-specific tuning through `params`. Bundled edges are faint by default,
so overlapping trunks read as density.
