---
title: "Sankey diagram"
description: "A multi-stage energy-flow diagram, with ribbons that fade from their source colour to their target."
categories: [Flows]
---

```{r}
#| include: false
source("_helper.R")
```

`vsankey()` builds a layered flow diagram from a *flow list* — one row per flow
with a `from` node, a `to` node, and a `value` (the ribbon width). Nodes are the
union of `from`/`to`, layered by longest path and ordered within each column to
minimise ribbon crossings.

```{r}
flows <- data.frame(
  from = c(
    "Coal", "Gas", "Nuclear", "Hydro", "Wind",
    "Grid", "Grid", "Grid",
    "Residential", "Industry", "Commercial"
  ),
  to = c(
    "Grid", "Grid", "Grid", "Grid", "Grid",
    "Residential", "Industry", "Commercial",
    "Losses", "Losses", "Losses"
  ),
  value = c(24, 18, 20, 8, 12, 32, 28, 22, 6, 9, 5)
)

p <- vsankey(
  flows, from, to, value,
  show_values = TRUE,
  flow_color = "gradient"
) |>
  labs(title = "Energy supply to end use")
```

Setting `flow_color = "gradient"` fades each ribbon from its source node's colour
to its target's, so a flow's origin and destination both read at a glance.

```{r}
#| echo: false
vg(p, "sankey")
```
